---
title: "A first try"
author: "Franzi"
date: "2 7 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Einlesen:

```{r}
train <- read.csv("~/R/Titanic/train.csv")
test <- read.csv("~/R/Titanic/test.csv")
library(dplyr)
library(Hmisc)
library(caret)
library(mlr)
library(tuneRanger)
```

## Explorative Datenanalyse

First glimpse

```{r}
head(train)
summary(train)
```

There are quite a number of na's in the age variable:

```{r}
sum(is.na(train$Age))
```

How to impute it? If there is more than 1 spouse/sibling around, it is likely that it is a child. Hence, we put the age to be the mean of all people with many siblings. If the person is female and the name includes "Mrs/Mlle", then she is (supposedly) married and we put the age the mean of all wives.



```{r}
meanAgeWithManySiblings <- mean(filter(train,!is.na(Age) & SibSp>1)$Age)
meanAgeofWifes <- mean(filter(train,!is.na(Age) & grepl("Mrs|Mlle",Name))$Age);
train %>% mutate(Age=ifelse(is.na(Age) & SibSp>1,meanAgeWithManySiblings,Age)) -> train
train %>% mutate(Age=ifelse(is.na(Age) & grepl("Mrs|Mlle",Name),meanAgeofWifes,Age)) -> train

test %>% mutate(Age=ifelse(is.na(Age) & SibSp>1,meanAgeWithManySiblings,Age)) -> test
test %>% mutate(Age=ifelse(is.na(Age) & grepl(c("Mlle"),Name),meanAgeofWifes,Age)) -> test

```

Impute fares

```{r}

meanFare <- function(pc,place){
                    mean(filter(train,Pclass==pc & Embarked == place)$Fare)
}
train %>% rowwise() %>% mutate(Fare=ifelse(is.na(Fare) | Fare=="0",meanFare(Pclass,Embarked),Fare)) -> train
test %>%  rowwise() %>% mutate(Fare=ifelse(is.na(Fare) | Fare == "0",meanFare(Pclass,Embarked),Fare)) -> test 
```

## Using Caret

Preprocessing

```{r}
trainsub <- as.data.frame(select(train, - (Name | Ticket | Cabin)))
trainsub$Survived <- as.factor(trainsub$Survived)

tit_train_control <- trainControl(method="cv",number=10)
preProcessing <- preProcess(select(trainsub,- (Survived)),method=c("knnImpute"))

trainPrepro <- predict(preProcessing,trainsub)

trainPrepro %>% rowwise() %>% mutate(Sex=ifelse(Sex=="female",0,1)) -> trainPrepro
trainPrepro %>% rowwise() %>% mutate(Embarked=case_when(Embarked == "S" ~ 0, Embarked == "Q" ~ 1, Embarked == "C" ~ 2, TRUE ~ 0)) -> trainPrepro
```

Entscheidungsbaum:

```{r}

my_tuning_grid <- expand.grid(cp = seq(0.01,0.08,0.01))

decisiontree <- caret::train(Survived ~ ., data=trainPrepro, method="xgbDART", trControl=tit_train_control)
#treebag/rpart

prediction1 <- predict(decisiontree,trainPrepro)

confusionMatrix(prediction1,trainPrepro$Survived)

```

Random forest with rf:

```{r}

my_tuning_grid <- expand.grid(mtry = c(2:4))

rftree <- caret::train(Survived ~ ., data=trainPrepro, method="rf",trControl=tit_train_control, tuneGrid=my_tuning_grid, importance = TRUE)
varImp(rftree)

prediction <- predict(rftree,trainPrepro)

confusionMatrix(prediction,trainPrepro$Survived)

```


Random forest with ranger:

```{r}

my_tuning_grid <- expand.grid(mtry = c(1,2, 3),
                      splitrule = c("gini", "extratrees"),
                      min.node.size = c(1, 3,5))

rangertree <- caret::train(Survived ~ ., data=trainPrepro, method="ranger",trControl=tit_train_control, tuneGrid=my_tuning_grid)

prediction <- predict(rangertree,trainPrepro)

confusionMatrix(prediction,trainPrepro$Survived)

```

Gradient boosting:

```{r}
myGrid <- expand.grid(n.trees = c(150, 200, 250, 300),
 interaction.depth = c(1, 2,3),
 shrinkage = c(0.05, 0.1, 0.15),
 n.minobsinnode = c(5,10))

gbm_tree <- caret::train(Survived ~., data = trainPrepro, method = "gbm", trControl = tit_train_control, verbose = FALSE, tuneGrid=myGrid)

prediction <- predict(gbm_tree,trainPrepro)

confusionMatrix(prediction,trainPrepro$Survived)
```


Compare models

```{r}
resamps <- resamples(list(gbm = gbm_tree, randomForest = rftree, dt=decisiontree,ranger=rangertree))
summary(resamps)


```
#Tune Ranger

```{r}
survival.task <- makeClassifTask(data=trainPrepro,target="Survived")
estimateTimeTuneRanger(survival.task)

res = tuneRanger(survival.task, num.trees = 1000, num.threads = 2, iters = 70, save.file.path = NULL)
res$model

tunerprediction <- predict(res$model, newdata = testPrepro)
#confusionMatrix(tunerprediction$data$response,trainPrePro$Survived)

```


Apply to test data:

```{r}
testsub <- select(test, - (Name | Ticket | Cabin))
testPrepro <- predict(preProcessing,testsub)
testPrepro %>% rowwise() %>% mutate(Sex=ifelse(Sex=="female",0,1)) -> testPrepro
testPrepro %>% rowwise() %>% mutate(Embarked=case_when(Embarked == "S" ~ 0, Embarked == "Q" ~ 1, Embarked == "C" ~ 2, TRUE ~ 0)) -> testPrepro
dtprediction <- predict(decisiontree,newdata=testPrepro,na.action = na.pass)
rfprediction <- predict(rftree,newdata=testPrepro,na.action = na.pass)
rangerprediction <- predict(rangertree,newdata=testPrepro,na.action = na.pass)
gbmprediction <- predict(gbm_tree,newdata=testPrepro)

#mixed model
 prediction<- data.frame(dt=as.integer(as.character(dtprediction)),rf=as.integer(as.character(rfprediction)),ranger=as.integer(as.character(rangerprediction)),gbm=as.integer(as.character(gbmprediction)))
  
prediction %>% rowwise() %>% mutate(decision=ifelse(dt+rf+ranger+gbm>=2,1,0)) -> prediction
  
results <- data.frame(PassengerId=testsub$PassengerId,Survived=prediction$decision)

#single model

#results <- data.frame(PassengerId=testsub$PassengerId,Survived=tunerprediction$data$response)


write.table(results,"~/R/Titanic/results.csv",sep=",",row.names = FALSE)

```
