---
title: "A first try"
author: "Franzi"
date: "2 7 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Einlesen:

```{r}
train <- read.csv("~/R/Titanic/train.csv")
test <- read.csv("~/R/Titanic/test.csv")
library(dplyr)
library(Hmisc)
library(caret)
```

## Explorative Datenanalyse

First glimpse

```{r}
head(train)
summary(train)
```

There are quite a number of na's in the age variable:

```{r}
sum(is.na(train$Age))
```

How to impute it? If there is more than 1 spouse/sibling around, it is likely that it is a child. Hence, we put the age to be the mean of all people with many siblings. If the person is female and the name includes "Mrs/Mlle", then she is (supposedly) married and we put the age the mean of all wives.



```{r}
meanAgeWithManySiblings <- mean(filter(train,!is.na(Age) & SibSp>1)$Age)
meanAgeofWifes <- mean(filter(train,!is.na(Age) & grepl("Mrs|Mlle",Name))$Age);
train %>% mutate(Age=ifelse(is.na(Age) & SibSp>1,meanAgeWithManySiblings,Age)) -> train
train %>% mutate(Age=ifelse(is.na(Age) & grepl("Mrs|Mlle",Name),meanAgeofWifes,Age)) -> train

test %>% mutate(Age=ifelse(is.na(Age) & SibSp>1,meanAgeWithManySiblings,Age)) -> test
test %>% mutate(Age=ifelse(is.na(Age) & grepl(c("Mlle"),Name),meanAgeofWifes,Age)) -> test

```

Impute fares

```{r}

meanFare <- function(pc,place){
                    mean(filter(train,Pclass==pc & Embarked == place)$Fare)
}
train %>% rowwise() %>% mutate(Fare=ifelse(is.na(Fare) | Fare=="0",meanFare(Pclass,Embarked),Fare)) -> train
test %>%  rowwise() %>% mutate(Fare=ifelse(is.na(Fare) | Fare == "0",meanFare(Pclass,Embarked),Fare)) -> test 
```

## Using Caret: Age, Sex and Fares

Entscheidungsbaum:

```{r}

trainsub <- as.data.frame(select(train, PassengerId | Survived | Sex | Age | Fare | Embarked))
trainsub$Survived <- as.factor(trainsub$Survived)

tit_train_control <- trainControl(method="cv",number=10)
preProcessing <- preProcess(select(trainsub,- (Survived)),method=c("knnImpute"))

trainPrepro <- predict(preProcessing,trainsub)

my_tuning_grid <- expand.grid(cp = seq(0.01,0.08,0.01))

decisiontree <- train(Survived ~ Age + Fare + Sex + Embarked, data=trainPrepro, method="treebag",trControl=tit_train_control)

prediction <- predict(decisiontree,trainPrepro)

confusionMatrix(prediction1,trainPrepro$Survived)

```

Random forest with rf:

```{r}

trainsub <- select(train, PassengerId | Survived | Sex | Age | Fare | Embarked)
trainsub$Survived <- as.factor(trainsub$Survived)

tit_train_control <- trainControl(method="cv",number=10)
preProcessing <- preProcess(as.data.frame(select(trainsub,- Survived)),method=c("knnImpute"))
trainPrepro <- predict(preProcessing,trainsub)

my_tuning_grid <- expand.grid(mtry = c(2:4))

rftree <- train(Survived ~ Age + Fare + Sex + Embarked, data=trainPrepro, method="rf",trControl=tit_train_control, tuneGrid=my_tuning_grid, importance = TRUE)
varImp(rftree)

prediction <- predict(rftree,trainPrepro)

confusionMatrix(prediction,trainPrepro$Survived)

```


Random forest with ranger:

```{r}

trainsub <- select(train, PassengerId | Survived | Sex | Age | Fare | Embarked)
trainsub$Survived <- as.factor(trainsub$Survived)

tit_train_control <- trainControl(method="cv",number=10)
preProcessing <- preProcess(as.data.frame(select(trainsub,- Survived)),method=c("knnImpute"))
trainPrepro <- predict(preProcessing,trainsub)

my_tuning_grid <- expand.grid(mtry = c(2, 3),
                      splitrule = c("gini", "extratrees"),
                      min.node.size = c(1, 2))

rangertree <- train(Survived ~ Age + Fare + Sex + Embarked, data=trainPrepro, method="ranger",trControl=tit_train_control, tuneGrid=my_tuning_grid)

prediction <- predict(rangertree,trainPrepro)

confusionMatrix(prediction,trainPrepro$Survived)

```

Gradient boosting:

```{r}
myGrid <- expand.grid(n.trees = c(150, 200, 250, 300),
 interaction.depth = c(1, 2,3),
 shrinkage = c(0.05, 0.1, 0.15),
 n.minobsinnode = c(5,10))

gbm_tree <- train(Survived ~ Age + Fare + Sex + Embarked, data = trainPrepro, method = "gbm", trControl = tit_train_control, verbose = FALSE, tuneGrid=myGrid)

prediction <- predict(gbm_tree,trainPrepro)

confusionMatrix(prediction,trainPrepro$Survived)
```

Apply to test data:

```{r}
testsub <- select(test, PassengerId | Sex | Age | Fare | Embarked)
testPrepro <- predict(preProcessing,testsub)
dtprediction <- predict(decisiontree,newdata=testPrepro,na.action = na.pass)
rfprediction <- predict(rftree,newdata=testPrepro,na.action = na.pass)
rangerprediction <- predict(rangertree,newdata=testPrepro,na.action = na.pass)
gbmprediction <- predict(gbm_tree,newdata=testPrepro)


prediction<- data.frame(dt=as.integer(as.character(dtprediction)),rf=as.integer(as.character(rfprediction)),ranger=as.integer(as.character(rangerprediction)),gbm=as.integer(as.character(gbmprediction)))
 
prediction %>% rowwise() %>% mutate(decision=ifelse(dt+rf+ranger+gbm>=2,1,0)) -> prediction


results <- data.frame(PassengerId=testsub$PassengerId,Survived=prediction$decision)
write.table(results,"~/R/Titanic/results.csv",sep=",",row.names = FALSE)

```
