---
title: "A first try"
author: "Franzi"
date: "2 7 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Einlesen:

```{r}
train <- read.csv("~/R/Titanic/train.csv")
test <- read.csv("~/R/Titanic/test.csv")
library(dplyr)
library(Hmisc)
library(caret)
```

## Explorative Datenanalyse

First glimpse

```{r}
head(train)
summary(train)
```

There are quite a number of na's in the age variable:

```{r}
sum(is.na(train$Age))
```

How to impute it? If there is more than 1 spouse/sibling around, it is likely that it is a child. Hence, we put the age to be 10. If the person is female and the name includes "Mrs", then she is (supposedly) married and we put the age to be 40.

```{r}
train %>% mutate(Age=ifelse(is.na(Age) & SibSp>1,10,Age)) -> train
train %>% mutate(Age=ifelse(is.na(Age) & grepl("Mrs|Mlle",Name),40,Age)) -> train

test %>% mutate(Age=ifelse(is.na(Age) & SibSp>1,10,Age)) -> test
test %>% mutate(Age=ifelse(is.na(Age) & grepl(c("Mlle"),Name),40,Age)) -> test

```

Impute fares (fixed mistake compared to v2)

```{r}

meanFare <- function(pc){
                    mean(filter(train,Pclass==pc)$Fare)
}
train %>% rowwise() %>% mutate(Fare=ifelse(is.na(Fare) | Fare=="0",meanFare(Pclass),Fare)) -> train
test %>% rowwise() %>% mutate(Fare=ifelse(is.na(Fare) | Fare == "0",meanFare(Pclass),Fare)) -> test 
```


## Using Caret

Entscheidungsbaum:

```{r}

trainsub <- select(train, PassengerId | Survived | Sex | Age | Fare | Embarked)
trainsub$Survived <- as.factor(trainsub$Survived)

tit_train_control <- trainControl(method="cv",number=10)
preProcessing <- preProcess(select(trainsub,- (Survived)),method=c("knnImpute"))

trainPrepro <- predict(preProcessing,trainsub)

my_tuning_grid <- expand.grid(cp = seq(0.01,0.06,0.01))

tree <- train(Survived ~ Age + Fare + Sex, data=trainPrepro, method="rpart",trControl=tit_train_control, tuneGrid=my_tuning_grid)

prediction <- predict(tree,trainPrepro)

confusionMatrix(prediction,trainPrepro$Survived)

```


Random forest:

```{r}

trainsub<- select(train, PassengerId | Survived | Sex | Age | Fare | Embarked)
trainsub$Survived <- as.factor(trainsub$Survived)

tit_train_control <- trainControl(method="cv",number=10)
preProcessing <- preProcess(select(trainsub,- (Survived)),method=c("knnImpute"))
trainPrepro <- predict(preProcessing,trainsub)

my_tuning_grid <- expand.grid(mtry = c(2, 3),
                      splitrule = c("gini", "extratrees"),
                      min.node.size = c(1, 2))

tree <- train(Survived ~ Age + Fare + Sex + Embarked, data=trainPrepro, method="ranger",trControl=tit_train_control, tuneGrid=my_tuning_grid)

prediction <- predict(tree,trainPrepro)

confusionMatrix(prediction,trainPrepro$Survived)

```

Apply to test data:

```{r}
testsub <- select(test, PassengerId | Sex | Age | Fare | Embarked)
testPrepro <- predict(preProcessing,testsub)
testprediction <- predict(tree,newdata=testPrepro,na.action = na.pass)

results <- data.frame(PassengerId=testsub$PassengerId,Survived=as.integer(as.character(testprediction)))
write.table(results,"~/R/Titanic/results.csv",sep=",",row.names = FALSE)

```
